---
title: "TidyTemplate"
date: 2020-09-16
output: html_output
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2020-09-15")
tt
kids <- tt$kids

```

How many variables are there in this data set?
```{r}
kids %>%
  distinct(variable)
```

We might do a:

* Choropleth (color by state)
* Animated choropleth (showing change over time 1997-2016)
* What states had the fastest growth/shrinkage

For now going to focus on school spending.  It looks like New York cut spending in 2015 overall, whilst California and Texas continued to increase spending. This continues to be true at a per child basis, however relatively New York spends almost double what the other spends per child.
```{r}
kids %>%
  filter(variable =="PK12ed")%>%
  group_by(year)%>%
  summarise(total = sum(inf_adj))%>%
  ggplot(aes(year,total))+
  geom_line()+
  expand_limits(y=0)+
  scale_y_continuous(labels = dollar)

kids %>%
  filter(variable =="PK12ed", state %in% c("New York", "Texas", "California"))%>%
  group_by(year)%>%
  ggplot(aes(year,inf_adj, color = state))+
  geom_line()+
  expand_limits(y=0)+
  scale_y_continuous(labels = dollar)

kids %>%
  filter(variable =="PK12ed", state %in% c("New York", "Texas", "California"))%>%
  group_by(year)%>%
  ggplot(aes(year,inf_adj_perchild, color = state))+
  geom_line()+
  expand_limits(y=0)+
  scale_y_continuous(labels = dollar)

```

Learning some new code, what if we wanted to look at 20 unique states? (surprise it's not actually new code, we are just using the sample code to randomly pick 20 states out of the 51 unique states).

Now it looks like most states saw a decrease or plateau in spending in 2009, when the financial crisis really took hold in America (not 2008 like over here).
```{r}
kids %>%
  filter(variable =="PK12ed", state %in% sample(unique(state),20))%>%
  group_by(year)%>%
  ggplot(aes(year,inf_adj_perchild, color = state))+
  geom_line()+
  expand_limits(y=0)+
  scale_y_continuous(labels = dollar)

kids %>%
  filter(variable =="PK12ed", state %in% sample(unique(state),51))%>%
  mutate(state = fct_reorder(state, inf_adj_perchild,sum, .desc=TRUE))%>%
  group_by(year)%>%
  ggplot(aes(year,1000* inf_adj_perchild))+
  geom_line()+
  geom_vline(xintercept=2009, color = "red", lty = 2)+
  expand_limits(y=0)+
  scale_y_continuous(labels = dollar)+
  facet_wrap(~state, nrow = 7)+
  labs(x ="Year", y = "Inflation-adjusted spending per child")

```

What if we look at change over time, rather than absolute value?
```{r}
kids %>%
  filter(variable=="PK12ed")%>%
  group_by(state)%>%
  mutate(index = inf_adj_perchild/inf_adj_perchild[1])%>%
  ungroup()%>%
  mutate(state = fct_reorder(state, index,max, .desc=TRUE))%>%
  ggplot(aes(year,index))+
  geom_line()+
  geom_vline(xintercept=2009, color = "red", lty = 2)+
  geom_hline(yintercept = 1, color = "darkgray")+
  scale_y_continuous(labels = percent, breaks = c(0,1,2,3))+
  facet_wrap(~state, nrow = 7)+
  labs(x ="Year", y = "Normalized inflation adjusted spend per child")

plot_faceted <- function(tbl, y_axis) {
  tbl %>%
    mutate(state = fct_reorder(state, {{ y_axis }}, max, .desc=TRUE))%>%
    ggplot(aes(year, {{ y_axis }}))+
    geom_hline(yintercept = 0, color = "darkgray")+
    geom_line()+
    facet_wrap(~state)
}


plot_index_faceted <- function(tbl) {
  tbl %>%
    group_by(state,variable)%>%
    mutate(index = inf_adj_perchild/inf_adj_perchild[1])%>%
    ungroup()%>%
    plot_faceted(index)+
    scale_y_continuous(labels = percent)
}  


kids %>%
  filter(variable =="PK12ed") %>%
  plot_index_faceted()+
  geom_vline(xintercept=2009, color = "red", lty = 2)+
  labs(x ="Year", y = "Normalized inflation adjusted spend per child")
```

Lets look at a different variable, higher education spending.  This is much more variable, and maybe doesn't work for this visualisation?

In some states, there is no increase over time, whilst some are increasing.
```{r}
kids %>%
  filter(variable =="highered") %>%
  plot_index_faceted()+
  labs(x ="Year", y = "Normalized inflation adjusted spend per child")

kids %>%
  filter(variable =="highered") %>%
  plot_faceted()+
  labs(x = "year", y = "inflation adjusted spend per child")
```

# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}



```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}


  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```
